Nitrogen's form determines its fate in the North Pacific Subtropical Gyre during PARAGON I

### 2025-05-12 manuscript/manuscript.Rmd
Cleaning up data load things, placing in separate chunk

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
library(tidyverse)
library(ggh4x)
library(patchwork)
library(ggtext)
library(minpack.lm)
options(pillar.sigfig = 7)
options(pillar.print_min = 100)

parametadata <- read_csv("metadata/parametadata.csv") %>%
  # filter(IS_type=="noIS") %>%
  filter(polarity=="pos") %>%
  filter(samp_type=="Smp") %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
  arrange(amendment, depth, timepoint_num, startime) %>%
  mutate(timepoint=fct_inorder(timepoint))
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
  left_join(parametadata %>% distinct(filename, shortname))
all_stans <- read_csv("metadata/all_stans.csv")
stan_patched <- all_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))

@@

samp_areas <- parametadata %>%
  filter(samp_type=="Smp") %>%
  distinct(shortname, depth) %>%
  left_join(peak_areas) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  filter(!compound_name%in%bad_cmpds) %>%
  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
  select(-depth)
labeled_fracs <- samp_areas %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
  select(compound_name, iso_name, shortname, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name)
```

@@

```{r thaa data}
thaa_quant <- read_csv("quant_data.csv")
metathaa <- read_csv("metathaa.csv") %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve")))

all_thaa_plot <- thaa_quant %>%
  filter(!str_detect(iso_name, "2H")) %>%
  left_join(metathaa) %>%
  filter(samp_type=="Smp") %>%
  left_join(aa_ncs) %>%
  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
  mutate(n_N_unlab=n_N-n_N_lab) %>%
  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
  left_join(metathaa) %>%
  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"),
                          labels=c("Initial", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
  summarise(total_uM=mean(total_uM), .by=c(amendment, timepoint, name)) %>%
  mutate(name=factor(name, levels=c("nM_N_unlab", "nM_N_lab"), labels = c("Unlabeled", "Labeled"))) %>%
  ggplot(aes(x=amendment, y=total_uM, fill=name)) +
  geom_col(color="black", width=1) +
  theme_bw() +
  scale_x_discrete(expand = expansion()) +
  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
  scale_fill_manual(breaks = c("Unlabeled", "Labeled"), values=c("grey80", "grey20")) +
  labs(x=NULL, y="Total uM N in THAAs", fill=NULL)

>>>>

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
library(tidyverse)
library(ggh4x)
library(patchwork)
library(ggtext)
library(minpack.lm)
options(pillar.sigfig = 7)
options(pillar.print_min = 100)
```

```{r load data}
parametadata <- read_csv("metadata/parametadata.csv") %>%
  # filter(IS_type=="noIS") %>%
  filter(polarity=="pos") %>%
  filter(samp_type=="Smp") %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(amendment=factor(amendment, levels=c("Amm", "Nit", "Arg", "GMP"))) %>%
  arrange(amendment, depth, timepoint_num, startime) %>%
  mutate(timepoint=fct_inorder(timepoint))
peak_areas <- read_csv("targeted/all_peak_areas_w_isos.csv") %>%
  left_join(parametadata %>% distinct(filename, shortname))
all_stans <- read_csv("metadata/all_stans.csv")
stan_patched <- all_stans %>%
  filter(!compound_name=="beta-Glutamic acid") %>%
  mutate(compound_name=str_replace(compound_name, "L-Glutamic acid", "L-Glutamic acid/beta-Glutamic acid")) %>%
  filter(!compound_name=="L-Isoleucine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Leucine", "L-Leucine/L-Isoleucine")) %>%
  filter(!compound_name=="L-Homoserine") %>%
  mutate(compound_name=str_replace(compound_name, "L-Threonine", "L-Threonine/L-Homoserine")) %>%
  filter(!compound_name=="Allopurinol") %>%
  mutate(compound_name=str_replace(compound_name, "Hypoxanthine", "Hypoxanthine/Allopurinol"))

@@

samp_areas <- parametadata %>%
  filter(samp_type=="Smp") %>%
  distinct(shortname, depth) %>%
  left_join(peak_areas) %>%
  mutate(shortname=fct_inorder(shortname)) %>%
  filter(!compound_name%in%bad_cmpds) %>%
  filter(!(depth=="Deep" & compound_name%in%maa_names)) %>%
  select(-depth)
labeled_fracs <- samp_areas %>%
  mutate(labeled_frac=1-area/sum(area), .by = c("shortname", "compound_name")) %>%
  select(compound_name, iso_name, shortname, labeled_frac) %>%
  filter(str_detect(iso_name, "13C0, 15N0")) %>%
  select(-iso_name)

thaa_quant <- read_csv("quant_data.csv")
metathaa <- read_csv("metathaa.csv") %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(startime=factor(startime, levels=c("Morn", "Eve")))
```

@@

```{r thaa data}
all_thaa_plot <- thaa_quant %>%
  filter(!str_detect(iso_name, "2H")) %>%
  left_join(metathaa) %>%
  filter(samp_type=="Smp") %>%
  left_join(aa_ncs) %>%
  select(filename, compound_name, iso_name, env_conc_nm, n_C, n_N) %>%
  mutate(n_N_lab=as.numeric(str_extract(iso_name, "(?<=15N)\\d+"))) %>%
  mutate(n_N_unlab=n_N-n_N_lab) %>%
  mutate(nM_N_lab=env_conc_nm*n_N_lab, nM_N_unlab=env_conc_nm*n_N_unlab) %>%
  select(filename, compound_name, iso_name, nM_N_lab, nM_N_unlab) %>%
  pivot_longer(c(nM_N_lab, nM_N_unlab)) %>%
  summarise(total_uM=sum(value)/1000, .by=c(filename, name)) %>%
  left_join(metathaa) %>%
  mutate(amendment=ifelse(timepoint=="T0", "T0", amendment)) %>%
  mutate(amendment=factor(amendment, levels = c("T0", "Amm", "Nit", "Arg", "GMP"),
                          labels=c("Initial", "+Ammonia", "+Nitrate", "+Arginine", "+GMP"))) %>%
  summarise(total_uM=mean(total_uM), .by=c(amendment, timepoint, name)) %>%
  mutate(name=factor(name, levels=c("nM_N_unlab", "nM_N_lab"), labels = c("Unlabeled", "Labeled"))) %>%
  ggplot(aes(x=amendment, y=total_uM, fill=name)) +
  geom_col(color="black", width=1) +
  theme_bw() +
  scale_x_discrete(expand = expansion()) +
  scale_y_continuous(expand = expansion(mult=c(0, 0.05))) +
  scale_fill_manual(breaks = c("Unlabeled", "Labeled"), values=c("grey80", "grey20")) +
  labs(x=NULL, y="Total uM N in THAAs", fill=NULL)
